---
title: "Anti-PEG"
output:
  rmarkdown::html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    theme: flatly
    highlight: zenburn
date: "2022-08-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "../..")
```

## Load in libraries

```{r}
#-------------------------------------------------------------------------------
# ANALYSIS OF ADCC DATA
#-------------------------------------------------------------------------------
library(tidyverse)
library(lme4)
library(merTools)
library(emmeans)
library(writexl)
library(DT)
library(cowplot)
library(MuMIn)
source("scripts/utils/utils.R")
modernacolors()
```

## Read in dataset

```{r}
set.seed(100)
d <- readxl::read_xlsx("processed_data/antiPEG.xlsx") %>%
  mutate(
    interval = factor(interval, levels = c("PBS", 8,6,4,3,2,1,"Prime Only")),
    dose = factor(dose, levels = c("0ug", "1ug", "10ug")),
    sheet = paste0(sheet, " Concentration"),
    log2_titer = log2(antiPEG_titer),
    case = ifelse(dose == "0ug", "PBS", "Dosed"),
    case = factor(case, levels = c("PBS", "Dosed"))
  )
```


## Figure 6A-B: Anti-PEG Antibody Responses

```{r, fig.width=9.89, fig.height = 7.89}
# Visualize the full dataset ---------------------------------------------------
d_dotplot <- d %>%
  mutate(interval = factor(interval, levels = c("PBS", "Prime Only", 1, 2, 3, 4, 6, 8)))
p_raw <-  d_dotplot %>%
  ggplot(aes(
    x = interval,
    y = antiPEG_titer,
    color = interval,
    shape = dose,
    group = dose
  )) +
  geom_point(position = position_jitterdodge(jitter.width = 1)) +
  stat_summary(
    aes(group = interaction(interval, dose)),
    data = d_dotplot,
    geom = "errorbar",
    width = 0.5,
    position = position_dodge(width = 1),
    size = 0.5,
    inherit.aes = TRUE
  ) +
  stat_summary(
    aes(group = interaction(dose, interval)),
    data = d_dotplot,
    geom = "bar",
    position = position_dodge(width = 1),
    size = 1,
    inherit.aes = TRUE,
    fill = NA
  ) +
  scale_shape_manual(values = 15:17) +
  scale_color_manual(
    values = c(
      modernalightgray,
      modernadarkgray,
      modernalightblue,
      modernaorange,
      modernapurple,
      modernagreen,
      modernablue,
      modernared
    )
  ) +
  scale_y_continuous(trans = "log10",
                     breaks = c(10, 100, 1000, 10000, 100000)) +
  scale_x_discrete(labels = ~ ifelse(
    str_detect(.x, "[[:digit:]]"),
    paste0(.x, "-week\ninterval"),
    .x
  )) +
  labs(
    y = "Concentration (ng/mL)",
    x = "Dose interval",
    color = "Dose interval",
    shape = "Dose (ug)"
  ) +
  facet_wrap( ~ sheet, nrow = 2, scales = "free_y") +
  theme_bw()
print(p_raw)
```

```{r echo = FALSE}
ggsave("figures_tables/antiPEG/antiPEG_dotplots.png", width = 9.89, height = 7.81)
```

## Model fitting

```{r}
#-----------------------------------------------------------------------------
# Fit simple mixed models with random intercept for dose level
#   One model each for IgG, IgM
#-----------------------------------------------------------------------------
spl <- group_split(d, sheet) %>%
  set_names(map_chr(., ~ .x$sheet[1]))
f <- purrr::map(spl, ~ lmer(log2_titer ~ interval + (1 | dose), data = .x))
```


### Diagnostics

The fitted model for IgG has:

+ AIC = `r AIC(f[str_detect(names(f), "IgG")][[1]])`
+ BIC = `r BIC(f[str_detect(names(f), "IgG")][[1]])`
+ Conditional adjusted $R^2$ = `r MuMIn::r.squaredGLMM(f[str_detect(names(f), "IgG")][[1]])[,"R2c"]`

and the fitted model for IgM has:

+ AIC = `r AIC(f[str_detect(names(f), "IgM")][[1]])`
+ BIC = `r BIC(f[str_detect(names(f), "IgM")][[1]])`
+ Conditional adjusted $R^2$ = `r MuMIn::r.squaredGLMM(f[str_detect(names(f), "IgM")][[1]])[,"R2c"]`

where conditional adjusted $R^2$ is defined as the proportion of variance explained by both the fixed and random effects in the model. The low conditional adjusted $R^2$ arises from the fact that a large proportion of variation in the data comes from within-group animal-to-animal variability, as can be seen in the above dotplots of the raw data.

Model fits are summarized below.

```{r}
map(f, summary)
```
### Residual plots

```{r}
par(mfrow = c(2, 2))
plot(
  fitted(f[[1]]),
  residuals(f[[1]]),
  xlab = "Fitted values",
  ylab = "Residuals",
  main = names(f)[1]
)
abline(h = 0, col = "tomato2")
qqnorm(residuals(f[[1]]))
qqline(residuals(f[[1]]))

plot(
  fitted(f[[2]]),
  residuals(f[[2]]),
  xlab = "Fitted values",
  ylab = "Residuals",
  main = names(f)[2]
)
abline(h = 0, col = "tomato2")
qqnorm(residuals(f[[2]]))
qqline(residuals(f[[2]]))
```

## Model-based group-wise analyses

### Supplementary Figure 11: Statistical comparisons of anti-PEG IgG and IgM antibody responses between mRNA-1273 dosing intervals over control

```{r}
rg <- purrr::map(f, ref_grid)
contrasts <-
  purrr::map(rg, ~ contrast(.x, method = "dunnett", type = "response"))

p <- imap(
  contrasts,
  ~ plot(.x) +
    scale_x_continuous(
      breaks = log2(c(0.75, 1, 2, 4, 8)),
      labels = ~ 2 ^ .x) +
    scale_y_discrete(
      breaks = c(
        "8 - PBS",
        "6 - PBS",
        "4 - PBS",
        "3 - PBS",
        "2 - PBS",
        "1 - PBS",
        "Prime Only - PBS"
      ),
      labels = c(
        "8 / PBS",
        "6 / PBS",
        "4 / PBS",
        "3 / PBS",
        "2 / PBS",
        "1 / PBS",
        "Prime Only / PBS"
      )
    ) +
    geom_vline(aes(xintercept = 0), col = "tomato2", linetype = 2) +
    ggtitle(str_remove_all(.y, "anti-PEG | Concentration")) +
    labs(x = "Fold change", y = "Contrast") +
    theme_bw()
)
p_emgrid <- cowplot::plot_grid(plotlist = p,
                               nrow = 1,
                               ncol = 2)
print(p_emgrid)
```

```{r echo = FALSE}
cowplot::save_plot(plot = p_emgrid, filename = "figures_tables/antiPEG/antiPEG_contrasts.png", nrow = 1, ncol = 2)
```

### Supplementary Table 8: Anti-PEG *P*-value table for group mean comparisons

```{r}
#-------------------------------------------------------------------------------
# Tables of P-values from mean comparisons
#-------------------------------------------------------------------------------
antipeg_p_table <-
  imap_dfr(contrasts, ~ as_tibble(.x) %>% mutate(Antibody = .y)) %>%
  mutate(
    "Pairwise comparison" = str_replace(contrast, pattern = " - ", replacement = " / "),
    .before = 1
  ) %>%
  dplyr::select(`Pairwise comparison`, Antibody, p.value) %>%
  mutate(Antibody = str_remove_all(Antibody, "anti-PEG | Concentration"), ) %>%
  rename("Adjusted P-value" = "p.value")
DT::datatable(antipeg_p_table)
```

```{r echo = FALSE}
writexl::write_xlsx(x = antipeg_p_table, path = "figures_tables/antiPEG/antipeg_pvals.xlsx")
```

## Session Information

```{r}
sessionInfo()
```