---
title: "ELISA"
date: "2022-08-15"
output:
  rmarkdown::html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    theme: flatly
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "../..")

```

## Load in libraries
```{r}
library(tidyverse)
library(readxl)
library(mgcv)
library(emmeans)
library(ggeffects)
library(lme4)
library(formattable)
library(ggpubr)
library(pzfx)
library(DT)
library(cowplot)
source("scripts/utils/utils.R")
```

## Load data
```{r}
# load data
d <- read_excel("processed_data/ELISA.xlsx")
```

## Observed data visualization

### Supplementary Figure 3. Impact of mRNA-1273 Prime-boost Interval on S2-P-specific Serum Binding IgG Antibody Titers Through 24 Weeks Post-boost
```{r, fig.width = 16, fig.height = 16}
#-------------------------------------------------------------------------------
# Supplementary Figure 1
#-------------------------------------------------------------------------------

# color panel
panel0 <-
  c("#c85979",
    "#6980ce",
    "#4aac8d",
    "#b460bd",
    "#cd5d39",
    "#b49541",
    "#4ab0aa")
dosepanel <- c("#cb6939", "#8264cb")
LLOQ <- 25  # lower limit of quantificatin


# trend over time
raw1 <-
d %>%
  filter(interval != "PBS") %>%
  mutate(interval = factor(interval)) %>%
  mutate(
    .,
    interval = forcats::fct_recode(
      interval,
      "8 weeks" = "8",
      "6 weeks" = "6",
      "4 weeks" = "4",
      "3 weeks" = "3",
      "2 weeks" = "2",
      "1 week" = "1",
      "prime only" = "0"
    )
  ) %>%
  mutate(., interval = factor(.$interval, levels = levels(.$interval)[8:1])) %>%
  mutate(
    day = forcats::fct_recode(
      factor(day),
      "Day 56 \n (Prime)" = "56",
      "Day 64 \n (1wk post-boost)" = "64",
      "Day 73 \n (2wk post-boost)" = "73",
      "Day 87 \n (4wk post-boost)" = "87",
      "Day 115 \n (8wk post-boost)" = "115",
      "Day 143 \n (12wk post-boost)" = "143",
      "Day 171 \n (16wk post-boost)" = "171",
      "Day 199 \n (20wk post-boost)" = "199",
      "Day 227 \n (24wk post-boost)" = "226"
    )
  ) %>%
  mutate(dose = forcats::fct_recode(factor(dose), "1ug" = "1", "10ug" = "10")) %>%
  ggplot(aes(
    x = factor(day),
    y = value,
    color = factor(interval)
  )) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_jitterdodge(jitter.width = .2)) +
  geom_hline(aes(yintercept = log10(LLOQ), lty = "LLOQ")) +
  scale_linetype_manual(values = 2) +
  facet_wrap(~ dose, nrow = 3) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  labs(
    x = " ",
    y = expression(log[10] ~ S2P ~ IgG ~ titer),
    color = "",
    linetype = ""
  ) +
  scale_color_manual(values = panel0) +
  scale_y_continuous(breaks = 1:7) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 16),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    strip.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16)
  )

# animal-level line plot
raw2 <-
d %>%
  filter(interval != "PBS") %>%
  mutate(
    interval = factor(interval),
    animal_id = paste("animal", animal_id),
    animal_id = factor(
      animal_id,
      levels = c(
        "animal 1",
        "animal 2",
        "animal 3",
        "animal 4",
        "animal 5",
        "animal 6",
        "animal 7",
        "animal 8",
        "animal 9",
        "animal 10"
      )
    )
  ) %>%
  mutate(
    .,
    interval = forcats::fct_recode(
      interval,
      "8 weeks" = "8",
      "6 weeks" = "6",
      "4 weeks" = "4",
      "3 weeks" = "3",
      "2 weeks" = "2",
      "1 weeks" = "1",
      "prime only" = "0"
    )
  ) %>%
  mutate(., interval = factor(.$interval, levels = levels(.$interval)[8:1])) %>%
  mutate(dose = forcats::fct_recode(factor(dose), "1ug" = "1", "10ug" = "10")) %>%
  ggplot(aes(
    x = day,
    y = value,
    color = factor(animal_id)
  )) +
  geom_line() +
  geom_jitter(position = position_jitterdodge(jitter.width = .2)) +
  geom_hline(aes(yintercept = log10(LLOQ), lty = "LLOQ")) +
  scale_linetype_manual(values = 2) +
  facet_grid(dose ~ interval) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  labs(
    x = " ",
    y = expression(log[10] ~ S2P ~ IgG ~ titer),
    color = "",
    linetype = ""
  ) +
  scale_color_manual(
    values = c(
      "#c25dba",
      "#72b24a",
      "#7561cf",
      "#ce9b44",
      "#807dc3",
      "#7a7d37",
      "#4faad9",
      "#c95c3f",
      "#4bac88",
      "#c8577b"
    )
  ) +
  scale_y_continuous(breaks = 1:7) +
  scale_x_continuous(breaks = c(56  , 87, 115, 143, 171, 199, 227)) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 16),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    strip.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16)
  )

cowplot::plot_grid(plotlist = list( raw1, raw2 ), labels = c("A", "B"), nrow = 2, ncol = 1)

#ggsave(filename = "figures_tables/ELISA/plot_raw.png", width = 16, height=16)

```

## Model fitting

### Residual diagnostics when model day as a continuous variable
```{r, fig.width = 12, fig.height = 6}
#-------------------------------------------------------------------------------
# Fit the model
#-------------------------------------------------------------------------------
dtemp <- filter(d, interval != "PBS" & interval != "0") %>%
  mutate(
    dose = factor(dose),
    day = as.numeric(day),
    animal_id = factor(animal_id),
    interval = factor(interval)
  ) %>%
  mutate(
    animal_unique_id = paste0(interval, "_", animal_id),
    animal_unique_id = factor(animal_unique_id)
  )

# fit a gam model
gam_fit <-
  mgcv::gam(
    value ~ interval + dose + s(day, k = 9) + interval:day +
      interval:dose +  dose:day + dose:day:interval +
      s(interval, bs = "re") + s(interval, day, bs = "re") ,
    data = dtemp,
    method = 'REML'
  )


# Residual diagnostics figure
p <- assumption_check(gam_fit)
title <-
  ggdraw() + draw_label(
    "ELISA residual diagnostics when model day as a continuous variable",
    x = 0.03,
    hjust = 0
  )
plot_grid(
  title,
  p,
  ncol = 1,
  rel_heights = c(0.1, 1),
  align = "v"
)


# fit a gam linear model
gam_fit_linear <-
  mgcv::gam(
    value ~ interval + dose + day + interval:day +
      interval:dose +  dose:day + dose:day:interval +
      s(interval, bs = "re") + s(interval, day, bs = "re") ,
    data = dtemp,
    method = 'REML'
  )

# fit a gam model at df = 8
gam_fit_df8 <-
  mgcv::gam(
    value ~ interval + dose + s(day, k = 8) + interval:day +
      interval:dose +  dose:day + dose:day:interval +
      s(interval, bs = "re") + s(interval, day, bs = "re") ,
    data = dtemp,
    method = 'REML'
  )
```
### Diagnostics
The model without smoothing spline on `day` does not fit as good as the selected model, where the selected model has  AIC = `r round(AIC(gam_fit))` and  BIC = `r round(BIC(gam_fit))` and the model without smoothing spline has AIC = `r round(AIC(gam_fit_linear))` and BIC = `r round(BIC(gam_fit_linear))`.

The selected model has degrees of freedom equals 9 for the smoothing spline, which is the maximum degrees of freedom allowed given the unique covariate combinations. AIC and BIC tests suggested degree of freedom equals 9 has the best fit. The model has df = 8 smoothing spline has AIC = `r round(AIC(gam_fit_df8))` and BIC = `r round(BIC(gam_fit_df8))`. 

Therefore, we add smoothing spline with df = 9 to variable `day`.

```{r, fig.width = 12, fig.height = 6}
# Residual diagnostics figure
p <- assumption_check(gam_fit_linear)
title <-
  ggdraw() + draw_label(
    "ELISA residual diagnostics when model day as a continuous variable",
    x = 0.03,
    hjust = 0
  )
plot_grid(
  title,
  p,
  ncol = 1,
  rel_heights = c(0.1, 1),
  align = "v"
)


```


### Figure 2. S2-P-specific Serum Binding IgG Antibody Titers
```{r, fig.width=12, fig.height=8}
#-------------------------------------------------------------------------------
# Figure 2. Spike-specific Serum Binding IgG Antibody Titers
#-------------------------------------------------------------------------------
plot1 <- ggemmeans(gam_fit, terms = c("day", "interval", "dose"))
plot(plot1) +
  scale_color_manual(values = panel0) +
  scale_x_continuous(
    breaks = c(56 , 64, 73, 87, 115, 143, 171, 199, 226),
    labels = c(
      "56(prime)",
      "64(1wk post-boost)",
      "73(2wk post-boost)",
      "87(4wk post-boost)",
      "115(8wk post-boost)",
      "143(12wk post-boost)",
      "171(16wk post-boost)",
      "199(20wk post-boost)",
      "227(24wk post-boost)"
    )
  ) +
  labs(
    title = expression(Figure2A ~ Predicted ~ values ~ "for" ~ log[10] ~ S2P ~
                         IgG ~ titer),
    y = expression(log[10] ~ S2P ~ IgG ~ titer)
  ) +
  theme(panel.grid.major = element_line(colour = "grey100")) +
  theme(
    axis.text = element_text(size = 12),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    axis.title = element_text(size = 18),
    strip.text = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    plot.title = element_text(size = 20),
    legend.position = "bottom"
  ) +
  facet_grid( ~ facet) 
```

```{r, fig.width=12, fig.height=8}
# plot effect of dose treating dose as numerical variable
plot2 <-
  ggemmeans(
    gam_fit,
    terms = c("day", "dose", "interval"),
    condition = c("day", "interval")
  )
plot(plot2) +
  #scale_color_manual(values=dosepanel) +
  scale_x_continuous(
    breaks = c(56 , 64, 73, 87, 115, 143, 171, 199, 226),
    labels = c(
      "56(prime)",
      "64(1wk post-boost)",
      "73(2wk post-boost)",
      "87(4wk post-boost)",
      "115(8wk post-boost)",
      "143(12wk post-boost)",
      "171(16wk post-boost)",
      "199(20wk post-boost)",
      "226(24wk post-boost)"
    )
  ) +
  labs(
    title = expression(Figure2B ~ Predicted ~ values ~ "for" ~ log[10] ~ S2P ~
                         IgG ~ titer),
    y = expression(log[10] ~ S2P ~ IgG ~ titer)
  ) +
  theme(panel.grid.major = element_line(colour = "grey100")) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    vjust = 1
  ),
  legend.position = "bottom") +
  facet_wrap( ~ facet)
```

### Residual diagnostics when model day as a factor
```{r, fig.width=12, fig.height=6}
# fit the model for mean comparison
dtemp <- filter(d, interval != "PBS" & interval != "0") %>%
  mutate(
    dose = as.factor(dose),
    animal_id = factor(animal_id),
    interval = factor(interval),
    day = factor(day)
  ) %>%
  mutate(animal_id = paste0(interval, "_", animal_id)) %>%
  mutate(animal_id = paste0(animal_id , "_", dose))

lm_fit <-
  lmer(value ~ interval * dose * day + (1 | animal_id) , data = dtemp)
p2 <- assumption_check(lm_fit)


# Residual diagnostics figure
title <-
  ggdraw() + draw_label(
    "ELISA residual diagnostics when model day as a discrete variable",
    x = 0.03,
    hjust = 0
  )
plot_grid(
  title,
  p2,
  ncol = 1,
  rel_heights = c(0.1, 1),
  align = "v"
)

```

### Supplementary Figure 1. Predicted Fold-change of Spike-specific Serum Binding IgG Antibody Titers
```{r, fig.width=12, fig.height=8}
# The results indicate that the peak immune response/ titer levels are achieved at 2 weeks post prime
grid <- ref_grid(lm_fit, cov.keep = c("day", "dose", "interval"))
emm_fit_day <- emmeans(grid, ~ day |  interval + dose)

#-------------------------------------------------------------------------------
# Supplementary Figure 2. Predicted Fold-change of Spike-specific Serum Binding IgG Antibody Titers.
#-------------------------------------------------------------------------------

set.seed(101)
summary(
  contrast(
    emm_fit_day,
    list(
      "Day 73 vs Day 56 \n(2wk post-boost) vs (Prime)" = c(-1, 0, 1, 0, 0, 0, 0, 0, 0),
      "Day 227 vs Day 56 \n(24wk post-boost) vs (Prime)" = c(-1, 0, 0, 0, 0, 0, 0, 0, 1)
    )
  ),
  infer = TRUE,
  adjust = "mvt",
  level = .95
)  %>%
  as_tibble() %>%
  mutate(
    .,
    interval = forcats::fct_recode(
      interval,
      "8 weeks" = "8",
      "6 weeks" = "6",
      "4 weeks" = "4",
      "3 weeks" = "3",
      "2 weeks" = "2",
      "1 weeks" = "1"
    )
  ) %>%
  mutate(., interval = factor(.$interval, levels = levels(.$interval)[8:1])) %>%
  mutate(dose = forcats::fct_recode(factor(dose), "1ug" = "1", "10ug" = "10")) %>%
  ggplot(aes(x = contrast , y = estimate, colour = interval)) +
  geom_point(position = position_dodge(width = 0.6), size = 2)  +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = .1,
    position = position_dodge(width = 0.6),
    size = 1
  ) +
  scale_color_manual(values = panel0) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  scale_y_continuous(breaks = log10(c(0.5, 1, 3, 10, 30, 100)), labels = c(0.5, 1, 3, 10, 30, 100)) +
  geom_hline(yintercept = 0,
             color = "darkcyan",
             lty = "dashed") +
  labs(x = "",
       y = expression(Estimated ~ fold ~ change ~ of ~ S2P ~ IgG ~ titer)) +
  facet_grid( ~ dose) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 18),
    strip.text = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    plot.title = element_text(size = 20),
    legend.position = "bottom"
  )

```

### Supplementary Table 1. Statistical comparisons of S2-P IgG antibody titer in mice between different mRNA-1273 dosing intervals.
```{r}
#-------------------------------------------------------------------------------
# Supplementary Table 1
#-------------------------------------------------------------------------------
grid <- ref_grid(gam_fit, cov.keep = c("day", "dose", "interval"))
em_fit <- emmeans::emmeans(grid, ~ interval | day + dose)
set.seed(100)
sum_fit <-
  summary(
    pairs(em_fit , reverse = TRUE),
    infer = TRUE,
    adjust = "mvt",
    level = .95
  )


# generate tables
sum_fit %>%
  as_tibble() %>%
  dplyr::select(contrast, estimate, day, dose, p.value) %>%
  mutate(estimate = 10 ^ estimate) %>%
  mutate(contrast = str_replace(contrast, " - ", " / ")) %>%
  rename(
    "Pairwise comparison" = contrast,
    "Fold change" = "estimate",
    "Adjusted P-value" = "p.value"
  ) %>%
  filter(`Adjusted P-value`  < 0.05) %>%
  DT::datatable(caption = "Statistical comparisons of S2-P IgG antibody titer in mice between different mRNA-1273 dosing intervals") %>%
  formatSignif(columns = c('Fold change', 'Adjusted P-value'),
               digits = 7)
```

### Supplementary Figure 4. Statistical comparisons of S2-P IgG Antibody Titers between different mRNA-1273 dosing intervals(Supplementary Table 1 visualization)
```{r, fig.width = 12, fig.height = 18}
sum_fit %>%
  as_tibble() %>%
  mutate(contrast = str_replace(contrast, " - ", " / ")) %>%
  mutate(day = as.factor(day)) %>%
  mutate(
    day = forcats::fct_recode(
      factor(day),
      "Day 56 \n(Prime)" = "56",
      "Day 64 \n(1wk post-boost)" = "64",
      "Day 73 \n(2wk post-boost)" = "73",
      "Day 87 \n(4wk post-boost)" = "87",
      "Day 115 \n(8wk post-boost)" = "115",
      "Day 143 \n(12wk post-boost)" = "143",
      "Day 171 \n(16wk post-boost)" = "171",
      "Day 199 \n(20wk post-boost)" = "199",
      "Day 227 \n(24wk post-boost)" = "226"
    )
  ) %>%
  rename(
    "Pairwise comparison" = contrast,
    "Fold change" = "estimate",
    "Adjusted P-value" = "p.value",
    "Day" = "day"
  ) %>%
  mutate(`Pairwise comparison` = str_remove_all(`Pairwise comparison` , "interval")) %>%
  mutate(dose = forcats::fct_recode(factor(dose), "1ug" = "1", "10ug" = "10")) %>%
  ggplot(aes(x = `Pairwise comparison` , y = `Fold change`, color = Day)) +
  geom_point(position = position_dodge(width = 2), size = 3)  +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = .1,
    position = position_dodge(width = 2),
    size = 1
  ) +
  scale_color_manual(
    values = c(
       "#ABCFE5",
      "#93C4DE",
      "#78B5D8",
      "#60A6D1",
      "#4A97C9",
      "#3787C0",
      "#2575B7",
      "#1664AB",
      "#09539D",
      "#084185"
    )
  ) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  scale_y_continuous(breaks = log10(c(0.5, 1, 3, 10, 30, 100)), 
                     labels = c(0.5, 1, 3, 10, 30, 100)) +
  geom_hline(yintercept = 0,
             color = "darkcyan",
             lty = "dashed") +
  labs(x = "",
       y = expression(Estimated ~ fold ~ change ~ of ~ S2P ~ IgG ~ titer)) +
  facet_grid(`Pairwise comparison` ~ dose,
             scales = "free",
             space = "free") +
  coord_flip() +
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 18),
    strip.text = element_text(size = 18),
    legend.text = element_text(size = 16,
                               margin = margin(t = 10)),
    legend.title = element_text(size = 18),
    plot.title = element_text(size = 20),
    strip.text.y = element_blank(),
    legend.position = "right"
  )

ggsave(filename = "figures_tables/ELISA/ELISA_emmeans.png", width = 16, height=18)

```

## Session information

```{r}
sessionInfo()
```

