---
title: "ELISA"
date: "2022-08-15"
output:
  rmarkdown::html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    theme: flatly
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "../..")

```

## Load in libraries
```{r}
library(tidyverse)
library(readxl)
library(mgcv)
library(emmeans)
library(ggeffects)
library(lme4)
library(formattable)
library(ggpubr)
library(pzfx)
library(DT)
library(cowplot)
library(gratia)
library(cenGAM)
source("scripts/utils/utils.R")
modernacolors()
```

## Load data
```{r}
# load data
d <- read_excel("processed_data/ELISA.xlsx")
```

## Observed data visualization

### Supplementary Figure 3. Impact of mRNA-1273 Prime-boost Interval on S2-P-specific Serum Binding IgG Antibody Titers Through 24 Weeks Post-boost
```{r, fig.width = 16, fig.height = 16}
#-------------------------------------------------------------------------------
# Supplementary Figure 1
#-------------------------------------------------------------------------------

# color panel
panel0 <-
  c(
    modernadarkblue2,
    modernaorange,
    modernapurple,
    modernagreen,
    modernablue,
    modernared,
    modernadarkgray
  )
dosepanel <- c("#cb6939", "#8264cb")
LLOQ <- 25  # lower limit of quantificatin


# trend over time
raw1 <-
  d %>%
  filter(interval != "PBS") %>%
  mutate(interval = factor(interval)) %>%
  mutate(
    .,
    interval = forcats::fct_recode(
      interval,
      "8-week interval" = "8",
      "6-week interval" = "6",
      "4-week interval" = "4",
      "3-week interval" = "3",
      "2-week interval" = "2",
      "1-week interval" = "1",
      "Prime only" = "0"
    )
  ) %>%
  mutate(., interval = factor(.$interval, levels = levels(.$interval)[8:1])) %>%
  mutate(
    day = forcats::fct_recode(
      factor(day),
      "Prime only" = "56",
      "1 wk post-boost" = "64",
      "2 wk post-boost" = "73",
      "4 wk post-boost" = "87",
      "8 wk post-boost" = "115",
      "12 wk post-boost" = "143",
      "16 wk post-boost" = "171",
      "20 wk post-boost" = "199",
      "24 wk post-boost" = "226"
    )
  ) %>%
  mutate(dose = forcats::fct_recode(
    factor(dose),
    "mRNA-1273 1 µg" = "1",
    "mRNA-1273 10 µg" = "10"
  )) %>%
  ggplot(aes(
    x = factor(day),
    y = value,
    color = factor(interval)
  )) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_jitterdodge(jitter.width = .2)) +
  geom_hline(aes(yintercept = log10(LLOQ), lty = "LLOQ")) +
  scale_linetype_manual(values = 2) +
  facet_wrap( ~ dose, nrow = 3) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  labs(
    x = " ",
    y = expression(log[10] ~ S2P ~ IgG ~ titer),
    color = "",
    linetype = ""
  ) +
  scale_color_manual(
    values =   c(
      modernared,
      modernablue,
      modernagreen,
      modernapurple,
      modernaorange,
      modernadarkblue2,
      modernadarkgray
    )
  ) +
  scale_y_continuous(breaks = 1:7) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 16),
    strip.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16)
  )

# animal-level line plot
raw2 <-
  d %>%
  filter(interval != "PBS") %>%
  mutate(
    interval = factor(interval),
    animal_id = paste("animal", animal_id),
    animal_id = factor(
      animal_id,
      levels = c(
        "animal 1",
        "animal 2",
        "animal 3",
        "animal 4",
        "animal 5",
        "animal 6",
        "animal 7",
        "animal 8",
        "animal 9",
        "animal 10"
      )
    )
  ) %>%
  mutate(
    .,
    interval = forcats::fct_recode(
      interval,
      "8-week interval" = "8",
      "6-week interval" = "6",
      "4-week interval" = "4",
      "3-week interval" = "3",
      "2-week interval" = "2",
      "1-week interval" = "1",
      "Prime only" = "0"
    )
  ) %>%
  mutate(., interval = factor(.$interval, levels = levels(.$interval)[8:1])) %>%
  mutate(dose = forcats::fct_recode(
    factor(dose),
    "mRNA-1273 1 µg" = "1",
    "mRNA-1273 10 µg" = "10"
  )) %>%
  ggplot(aes(
    x = day,
    y = value,
    color = factor(animal_id)
  )) +
  geom_line() +
  geom_jitter(position = position_jitterdodge(jitter.width = .2)) +
  geom_hline(aes(yintercept = log10(LLOQ), lty = "LLOQ")) +
  scale_linetype_manual(values = 2) +
  facet_grid(dose ~ interval) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  labs(
    x = " ",
    y = expression(log[10] ~ S2P ~ IgG ~ titer),
    color = "",
    linetype = ""
  ) +
  scale_color_manual(
    values = c(
      "#c25dba",
      "#72b24a",
      "#7561cf",
      "#ce9b44",
      "#807dc3",
      "#7a7d37",
      "#4faad9",
      "#c95c3f",
      "#4bac88",
      "#c8577b"
    )
  ) +
  scale_y_continuous(breaks = 1:7) +
  scale_x_continuous(
    breaks = c(56  , 87, 115, 143, 171, 199, 227),
    labels = c(
      "Prime only",
      "4 wk post-boost",
      "8 wk post-boost",
      "12 wk post-boost",
      "16 wk post-boost",
      "20 wk post-boost",
      "24 wk post-boost"
    )
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 16),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    strip.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16)
  )

cowplot::plot_grid(
  plotlist = list(raw1, raw2),
  labels = c("A", "B"),
  nrow = 2,
  ncol = 1
)
```
```{r echo = FALSE}
ggsave(filename = "figures_tables/ELISA/plot_raw.png", width = 18, height=16)
ggsave(filename = "figures_tables/ELISA/plot_raw.pdf", width = 18, height=16)
```

## Model fitting
### Model selection

We use a generalized additive model (GAM) with a 9-dimensional thin plate spline basis in day. We try 4 different GAMs with varying complexity.

  + Model A: a GAM censored regression model that accounts for censored values in prime-only group, dosing interval group-specific smooth nonlinear trend in days, a dose-specific smooth nonlinear trend in days, animal-specific random effect, and animal_specific effect of day

  + Model B: a GAM regression model that exclude the prime-only group, with dosing interval group-specific smooth nonlinear trend in days, a dose-specific smooth nonlinear trend in days, animal-specific random effect, and animal_specific effect of day

  + Model C: a GAM regression model that exclude the prime-only group, with animal-specific random effect, and animal_specific effect of day

  + Model D: a GAM regression model that exclude the prime-only group, with a single day-specific smooth, dosing interval specific random effect, and dosing interval specific effect of day

Model B is the best model among the four choices. 
  
  + Model A does not fit as good as Model B which excluded the prime-only data. Note prime-only data has a different trend as other dosing interval groups with some censored observations. 
  + Model C that is without smoothing splines does not fit as good as Model B that has smoothing splines. 
  + Model B's residual diagnostics look slightly better than Model D. Model D has smaller `BIC` but larger `AIC` than Model B. Based on `AIC` and residual diagnostics, Model B is better than Model D.


```{r}
#-------------------------------------------------------------------------------
# Fit the model
#-------------------------------------------------------------------------------
dtemp <- filter(d, interval != "PBS") %>%
  mutate(
    dose = factor(dose),
    day = as.numeric(day),
    animal_id = factor(animal_id),
    interval = factor(interval)
  ) %>%
  mutate(
    animal_unique_id = paste0(interval, "_", animal_id),
    animal_unique_id = factor(animal_unique_id)
  )

# fit a gam censored model that includes prime-only group with dosing interval group-specific smooth nonlinear trend in days, a dose-specific smooth nonlinear trend in days, animal-specific random effect, and animal_specific effect of day
gam_fit_A <-
  mgcv::gam(
    value ~ interval + dose + interval:day +
      interval:dose +  dose:day + dose:day:interval +
      s(day, by = dose, bs="tp", k = 9) + s(day, by = interval,bs="tp", k = 9) +
      s(animal_unique_id, bs = "re") + s(animal_unique_id, day, bs = "re"),
    data = dtemp,
    family = tobit1(left.threshold = 1.097),
    method = 'REML'
  )

# fit a gam model the same as previous model but without prime-only group (interval = "0") 
d_mod <-  dtemp %>%
  filter(interval != "0")%>%
  droplevels()

gam_fit_B<-
  mgcv::gam(
    value ~ interval + dose + interval:day +
      interval:dose +  dose:day + dose:day:interval +
      s(day, by = dose, bs="tp", k = 9) + s(day, by = interval, bs="tp", k = 9) +
      s(animal_unique_id, bs = "re") + s(animal_unique_id, day, bs = "re"),
    data = d_mod,
    method = 'REML'
  )

# fit a gam model without prime-only group (interval = "0") with animal-specific random effect, and animal_specific effect of day
gam_fit_C <-
  mgcv::gam(
    value ~ interval + dose + day + interval:day +
      interval:dose +  dose:day + dose:day:interval +
      s(animal_unique_id, bs = "re") + s(animal_unique_id, day, bs = "re"),
    data = d_mod,
    method = 'REML'
  )

# fit a gam model without prime-only group (interval = "0") with a single day-specific smooth, dosing interval specific random effect, and dosing interval specific effect of day
gam_fit_D <-
  mgcv::gam(
    value ~ interval + dose + s(day, k = 9) + interval:day +
      interval:dose +  dose:day + dose:day:interval +
      s(interval, bs = "re") + s(interval, day, bs = "re") ,
    data = d_mod,
    method = 'REML'
  )

```

#### GAMs
##### Estimated smooths visualization

```{r, fig.width = 12, fig.height = 8}
plot_grid(gratia::draw(gam_fit_A),  labels = c("A"), vjust = 1) 
plot_grid(gratia::draw(gam_fit_B),  labels = c("B"), vjust = 1) 
plot_grid(gratia::draw(gam_fit_C),  labels = c("C"), vjust = 1) 
plot_grid(gratia::draw(gam_fit_D),  labels = c("D"), vjust = 1) 
```

##### Residual diagnostics
```{r, fig.width = 12, fig.height = 8}
# Residual diagnostics plots
list(gam_fit_A, gam_fit_B, gam_fit_C, gam_fit_D) %>%
  map(~assumption_check(.x, .x$model$value))%>%
  cowplot::plot_grid(plotlist = ., nrow = 2, labels = c("A", "B", "C", "D"))

```

##### AIC and BIC comparison
```{r}
AIC(gam_fit_A, gam_fit_B, gam_fit_C, gam_fit_D)%>% 
  rownames_to_column("Model") %>% 
  DT::datatable()%>%
  formatSignif(columns = c( 'df', 'AIC'),
               digits = 6)

BIC(gam_fit_A, gam_fit_B, gam_fit_C, gam_fit_D)%>% 
  rownames_to_column("Model") %>% 
  DT::datatable() %>%
  formatSignif(columns = c( 'df', 'BIC'),
               digits = 6)
```

#### Selected model residual diagnostics
```{r}
# Residual diagnostics figure
gam_fit <-  gam_fit_B
p <- assumption_check(gam_fit, gam_fit$model$value)
title <-
  ggdraw() + draw_label(
    "ELISA residual diagnostics when model day as a continuous variable",
    x = 0.03,
    hjust = 0
  )
plot_grid(
  title,
  p,
  ncol = 1,
  rel_heights = c(0.1, 1),
  align = "v"
)

```

### Residual diagnostics
### Figure 2. S2-P-specific Serum Binding IgG Antibody Titers
```{r, fig.width=12, fig.height=8}

#-------------------------------------------------------------------------------
# Figure 2. Spike-specific Serum Binding IgG Antibody Titers
#-------------------------------------------------------------------------------
plot1 <- ggemmeans(gam_fit, terms = c("day", "interval", "dose")) 

F2A <-
plot(plot1) +
  scale_color_manual(values = panel0, 
                     labels = c("1-week interval", "2-week interval", 
                                "3-week interval", "4-week interval", 
                                "6-week interval", "8-week interval")) +
  scale_fill_manual(values = panel0) +
  scale_x_continuous(
    breaks = c(56 , 64, 73, 87, 115, 143, 171, 199, 226),
    labels = c(
      "Prime only",
      "1 wk post-boost",
      "2 wk post-boost",
      "4 wk post-boost",
      "8 wk post-boost",
      "12 wk post-boost",
      "16 wk post-boost",
      "20 wk post-boost",
      "24 wk post-boost"
    )
  ) +
  labs(
    title = "", color = "", x = "",
    y = expression(log[10] ~ S2P ~ IgG ~ titer)
  ) +
  theme(panel.grid.major = element_line(colour = "grey100")) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    legend.position = "bottom"
  ) +
  facet_grid(~ facet, labeller = labeller(.cols = ~ paste("mRNA-1273", .x, "µg")))
```

```{r, fig.width=10, fig.height=12}
# plot effect of dose treating dose as numerical variable
plot2 <-
  ggemmeans(
    gam_fit,
    terms = c("day", "dose", "interval"),
    condition = c("day", "interval")
  )

F2B <-
plot(plot2) +
  scale_color_manual(values = c(modernared, modernablue),
                     labels = c("mRNA-1273 1 µg", "mRNA-1273 2 µg")) +
  scale_x_continuous(
    breaks = c(56 , 64, 73, 87, 115, 143, 171, 199, 226),
    labels = c(
      "Prime only",
      "1 wk post-boost",
      "2 wk post-boost",
      "4 wk post-boost",
      "8 wk post-boost",
      "12 wk post-boost",
      "16 wk post-boost",
      "20 wk post-boost",
      "24 wk post-boost"
    )
  ) +
  labs(
    x = "",
    title = "", color = "",
    y = expression(log[10] ~ S2P ~ IgG ~ titer)
  ) +
  theme(panel.grid.major = element_line(colour = "grey100")) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    vjust = 1
  ),
  legend.position = "bottom") +
  facet_wrap(~ facet, labeller = labeller(.cols = ~paste(.x, "week interval")))

Figure2 <-
cowplot::plot_grid(
  plotlist = list(F2A, F2B),
  labels = c("A", "B"),
  nrow = 2,
  ncol = 1
)

Figure2
```


```{r echo=FALSE}
ggsave(
  filename = "figures_tables/ELISA/ELISA_Titers_F2.pdf",
  plot = Figure2,
  width = 10,
  height = 12,
  device = cairo_pdf
)

ggsave(
  filename = "figures_tables/ELISA/ELISA_Titers_F2.png",
  plot = Figure2,
  width = 10,
  height = 12,
  bg = "white"
)

```

### Residual diagnostics when model day as a factor
```{r, fig.width=12, fig.height=6}
# fit the model for mean comparison
dtemp <- filter(d, interval != "PBS" & interval != "0") %>%
  mutate(
    dose = as.factor(dose),
    animal_id = factor(animal_id),
    interval = factor(interval),
    day = factor(day)
  ) %>%
  mutate(animal_id = paste0(interval, "_", animal_id)) %>%
  mutate(animal_id = paste0(animal_id , "_", dose))

lm_fit <-
  lmer(value ~ interval * dose * day + (1 |
                                          animal_id) , data = dtemp)
p2 <- assumption_check(lm_fit, lm_fit@frame$value)


# Residual diagnostics figure
title <-
  ggdraw() + draw_label(
    "ELISA residual diagnostics when model day as a discrete variable",
    x = 0.03,
    hjust = 0
  )
plot_grid(
  title,
  p2,
  ncol = 1,
  rel_heights = c(0.1, 1),
  align = "v"
)

```

### Supplementary Figure 1. Predicted Fold-change of Spike-specific Serum Binding IgG Antibody Titers
Each dot corresponds to the estimated fold change for the contrasted prime-boost dosing interval. Error bars are multivariate t adjusted 95% CIs on the estimated contrast.

```{r, fig.width=12, fig.height=8}
# The results indicate that the peak immune response/ titer levels are achieved at 2 weeks post prime
grid <- ref_grid(lm_fit, cov.keep = c("day", "dose", "interval"))
emm_fit_day <- emmeans(grid, ~ day |  interval + dose)

#-------------------------------------------------------------------------------
# Supplementary Figure 2. Predicted Fold-change of Spike-specific Serum Binding IgG Antibody Titers.
#-------------------------------------------------------------------------------

set.seed(101)
summary(
  contrast(
    emm_fit_day,
    list(
      "2 weeks post-boost \n vs Pre-boost" = c(-1, 0, 1, 0, 0, 0, 0, 0, 0),
      "24 weeks post-boost \n vs Pre-boost" = c(-1, 0, 0, 0, 0, 0, 0, 0, 1)
    )
  ),
  infer = TRUE,
  adjust = "mvt",
  level = .95
)  %>%
  as_tibble() %>%
  mutate(
    .,
    interval = forcats::fct_recode(
      interval,
      "8-week interval" = "8",
      "6-week interval" = "6",
      "4-week interval" = "4",
      "3-week interval" = "3",
      "2-week interval" = "2",
      "1-week interval" = "1"
    )
  ) %>%
  mutate(., interval = factor(.$interval, levels = levels(.$interval)[8:1])) %>%
  mutate(dose = forcats::fct_recode(
    factor(dose),
    "mRNA-1273 1 µg" = "1",
    "mRNA-1273 10 µg" = "10"
  )) %>%
  ggplot(aes(x = contrast , y = estimate, colour = interval)) +
  geom_point(position = position_dodge(width = 0.6), size = 2)  +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = .1,
    position = position_dodge(width = 0.6),
    size = 1
  ) +
  scale_color_manual(
    values =   c(
      modernared,
      modernablue,
      modernagreen,
      modernapurple,
      modernaorange,
      modernadarkblue2,
      modernadarkgray
    )
  ) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  scale_y_continuous(breaks = log10(c(0.5, 1, 3, 10, 30, 100)), labels = c(0.5, 1, 3, 10, 30, 100)) +
  geom_hline(yintercept = 0,
             color = "darkcyan",
             lty = "dashed") +
  labs(x = "",
       y = expression(Estimated ~ fold ~ change ~ of ~ S2P ~ IgG ~ titer)) +
  facet_grid(~ dose) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 18),
    strip.text = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    plot.title = element_text(size = 20),
    legend.position = "bottom"
  )
```

```{r echo = FALSE}
ggsave(filename = "figures_tables/ELISA/ELISA_foldchange_SF2.png",
       width = 10,
       height = 8)
ggsave(filename = "figures_tables/ELISA/ELISA_foldchange_SF2.pdf",
       width = 10,
       height = 8)


```

### Supplementary Table 1. Statistical comparisons of S2-P IgG antibody titer in mice between different mRNA-1273 dosing intervals.
```{r}
#-------------------------------------------------------------------------------
# Supplementary Table 1
#-------------------------------------------------------------------------------
grid <- ref_grid(gam_fit, cov.keep = c("day", "dose", "interval"))
em_fit <- emmeans::emmeans(grid, ~ interval | day + dose)
set.seed(100)
sum_fit <-
  summary(
    pairs(em_fit , reverse = TRUE),
    infer = TRUE,
    adjust = "mvt",
    level = .95
  )


# generate tables
elisa_p_table <-
sum_fit %>%
  as_tibble() %>%
  mutate(day = forcats::fct_recode(
      factor(day),
      "0 (pre-dose 2)" = "56",
      "7 (1)" = "64",
      "16 (2)" = "73",
      "30 (4)" = "87",
      "58 (8)" = "115",
      "86 (12)" = "143",
      "114 (16)" = "171",
      "142 (20)" = "199",
      "169 (24)" = "226"
    ))%>%
  dplyr::select(contrast, estimate, day, dose, p.value) %>%
  mutate(estimate = 10 ^ estimate) %>%
  mutate(contrast = str_remove_all(contrast, "interval"),
         contrast = str_replace(contrast, " - ", " / ")) %>%
  rename(
    "Pairwise comparison" = contrast,
    "Fold change" = "estimate",
    "Adjusted P-value" = "p.value"
  ) %>%
  mutate(`Adjusted P-value` = round(`Adjusted P-value`, digits = 6),
         `Fold change`= round(`Fold change`, digits = 6))  %>%
  filter(`Adjusted P-value`  < 0.05) 
  

elisa_p_table %>%
  DT::datatable(caption = "Statistical comparisons of S2-P IgG antibody titer in mice between different mRNA-1273 dosing intervals")


```

```{r echo = FALSE}
writexl::write_xlsx(elisa_p_table, "figures_tables/ELISA/ELISA_P.xlsx")
```

### Supplementary Figure 4. Statistical comparisons of S2-P IgG Antibody Titers between different mRNA-1273 dosing intervals(Supplementary Table 1 visualization)
Each dot corresponds to the estimated fold change for the contrasted prime-boost dosing interval. Error bars are multivariate t adjusted 95% CIs on the estimated contrast.

```{r, fig.width = 12, fig.height = 18}
sum_fit %>%
  as_tibble() %>%
  mutate(contrast  = str_remove_all(contrast  , "interval")) %>%
  mutate(contrast = str_replace(contrast, " - ", " weeks / ")) %>%
  mutate(contrast = ifelse(
    str_detect(contrast, "1"),
    paste(contrast, "week"),
    paste(contrast, "weeks")
  )) %>%
  mutate(day = as.factor(day)) %>%
  mutate(
    day = forcats::fct_recode(
      factor(day),
      "Prime only" = "56",
      "1 wk post-boost" = "64",
      "2 wk post-boost" = "73",
      "4 wk post-boost" = "87",
      "8 wk post-boost" = "115",
      "12 wk post-boost" = "143",
      "16 wk post-boost" = "171",
      "20 wk post-boost" = "199",
      "24 wk post-boost" = "226"
    )
  ) %>%
  rename(
    "Pairwise comparison" = contrast,
    "Fold change" = "estimate",
    "Adjusted P-value" = "p.value",
    "Day" = "day"
  ) %>%
  mutate(dose = forcats::fct_recode(
    factor(dose),
    "mRNA-1273 1 µg" = "1",
    "mRNA-1273 10 µg" = "10"
  )) %>%
  ggplot(aes(x = `Pairwise comparison` , y = `Fold change`, color = Day)) +
  geom_point(position = position_dodge(width = 2), size = 3)  +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = .1,
    position = position_dodge(width = 2),
    size = 1
  ) +
  scale_color_manual(
    values = c(
      "#ABCFE5",
      "#93C4DE",
      "#78B5D8",
      "#60A6D1",
      "#4A97C9",
      "#3787C0",
      "#2575B7",
      "#1664AB",
      "#09539D",
      "#084185"
    )
  ) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  scale_y_continuous(breaks = log10(c(0.5, 1, 3, 10, 30, 100)),
                     labels = c(0.5, 1, 3, 10, 30, 100)) +
  geom_hline(yintercept = 0,
             color = "darkcyan",
             lty = "dashed") +
  labs(x = "Prime-boost dosing interval",
       y = expression(Estimated ~ fold ~ change ~ of ~ S2P ~ IgG ~ titer)) +
  facet_grid(`Pairwise comparison` ~ dose,
             scales = "free",
             space = "free") +
  coord_flip() +
  theme(
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 18),
    strip.text = element_text(size = 18),
    legend.text = element_text(size = 16,
                               margin = margin(t = 10)),
    legend.title = element_text(size = 18),
    plot.title = element_text(size = 20),
    strip.text.y = element_blank(),
    legend.position = "right"
  )

```

```{r echo = FALSE}
ggsave(filename = "figures_tables/ELISA/ELISA_emmeans.png",
       width = 16,
       height = 18)
ggsave(filename = "figures_tables/ELISA/ELISA_emmeans.pdf",
       width = 16,
       height = 18)

```

## Session information

```{r}
sessionInfo()
```

