---
title: "ELISpot"
date: "2022-08-15"
output:
  rmarkdown::html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    theme: flatly
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "../..")

```

## Load in libraries
```{r}
library(tidyverse)
library(readxl)
library(mgcv)
library(emmeans)
library(ggeffects)
library(lme4)
library(DT)
library(ggpubr)
library(cowplot)
source("scripts/utils/utils.R")
```

## Load data
```{r}
# load ELISpot data
temp <- read_excel("processed_data/ELISpot.xlsx")

dp <- temp[!is.na(temp$counts), ] %>%
  filter(`low cell number (<0.15M)`== "no",
         day != "143")%>%
  mutate(counts = counts + 0.001)

d <-
  dp %>%
  filter(interval != "PBS")%>%
  mutate(animal_id = factor(animal_id),
         interval = factor(interval),
         day = factor(day))
```

## Observed data visualization

```{r, fig.width=12, fig.height=8}
dp %>%
  filter(interval != "PBS") %>%
  mutate(interval = factor(interval, levels = c("PBS", 0:8))) %>%
  mutate(
    interval = forcats::fct_recode(
      interval,
      "8 weeks" = "8",
      "6 weeks" = "6",
      "4 weeks" = "4",
      "3 weeks" = "3",
      "2 weeks" = "2",
      "1 week" = "1",
      "Prime only" = "0"
    )
  ) %>%
  mutate(day = forcats::fct_recode(factor(day), "day 64" = "64")) %>%
  ggplot(aes(x = interval, y = counts)) +
  geom_boxplot(outlier.shape = NA) +
  scale_shape_manual(values = c(16, 5)) +
  scale_y_log10(breaks = c(100, 500, 1000, 1500)) +
  geom_jitter(aes(x = interval, y = counts)) +
  stat_summary(
    fun.min = mean,
    fun = mean,
    fun.max = mean,
    color = "red",
    width = .5,
    geom = "errorbar",
    size = 1
  ) +
  facet_wrap(tissue ~ coating, scales = "free") +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  labs(
    x = "",
    y = "antigen specific \n ASC counts/million cells",
    color = "",
    linetype = "",
    title = "Antigen specific spleen and BM results"
  ) +
  theme(
    title = element_text(size = 14),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "bottom"
  )
```

## Model fitting for spleen data

```{r}
#-------------------------------------------------------------------------------
# Fit the model for spleen data 
#-------------------------------------------------------------------------------
# boxcox function to fit boxcox regression
boxcox_fit <- function(dat) {
  require(MASS)
  if (length(unique(dat$day)) == 1) {
    out <- boxcox(counts ~ interval, data = dat)
    optimal <- out$x[which.max(out$y)]
    bctran <- make.tran("boxcox", optimal)
    lm_fit <- with(bctran,
                   lm(linkfun(counts) ~ interval, data = dat))
  } else{
    out <- boxcox(counts ~ interval * day, data = dat)
    optimal <- out$x[which.max(out$y)]
    bctran <- make.tran("boxcox", optimal)
    lm_fit <- with(bctran,
                   lm(linkfun(counts) ~ interval * day, data = dat))
  }
  return(list(lm_fit = lm_fit, lambda = optimal))
}

# box cox model fit for spleen data
d_tissue <- subset(d, tissue == "spleen") %>%
  filter(!(coating == "NTD" & interval == "0"))

boxcox_fits <-
  d_tissue %>%
  split(.$coating) %>%
  map( ~ boxcox_fit(.x))

lm_fits <-
  purrr::transpose(boxcox_fits)$lm_fit

# obtain optimal lambda for boxcox transformation
lambdas <-
  purrr::transpose(boxcox_fits)$lambda %>% unlist()

assumption_plots <-
  imap(
    lm_fits ,
    ~ assumption_check(.x) %>%
      plot_grid(
        ggdraw() + draw_label(paste0("Spleen ", .y, "-specific model residual diagnostics "),
                              x = 0.03,
                              hjust = 0),
        .,
        ncol = 1,
        rel_heights = c(0.1, 1),
        align = "v"
      )
  )

# obtain emmeans
lm_grid <- ref_grid(lm_fits$S2P, cov.keep = c("interval", "day"))
emmt_dat <-
  emmeans(lm_grid, specs = ~ interval, type = "response") %>%
  as_tibble()

```

### Residual diagnostics for spleen data

```{r, fig.width = 12, fig.height = 6}
ggarrange(plotlist = assumption_plots)
```

### Figure 4A. Spike-specific Antibody Secreting Cells and Long-Lived Plasma Cells

```{r, fig.height=8}
#-------------------------------------------------------------------------------
# Figure 4A. Spike-specific Antibody Secreting Cells and Long-Lived Plasma Cells.
#-------------------------------------------------------------------------------
# fold change figure 
foldchange1 <-
  emmt_dat %>%
  mutate(
    interval = forcats::fct_recode(
      interval,
      "8 weeks" = "8",
      "6 weeks" = "6",
      "4 weeks" = "4",
      "3 weeks" = "3",
      "2 weeks" = "2",
      "1 week" = "1",
      "Prime only" = "0"
    )
  ) %>%
  ggplot(aes(x = interval , y = response)) +
  geom_point(position = position_dodge(width = 0.5), size = 2)  +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = .1,
    position = position_dodge(width = 0.5),
    size = 1
  ) +
  scale_color_manual(values = c(
    "#8d68ca",
    "#64a758",
    "#c85994",
    "#b6933f",
    "#609bce",
    "#ca5c49"
  )) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  scale_y_log10(breaks = c(100, 500, 1000, 1600)) +
  ggtitle("S2P specific spleen day 64") +
  labs(x = "", y = "Estimated ASC counts/million cells") +
  geom_hline(yintercept = emmt_dat$lower.CL[c(6, 7)],
             color = "darkcyan",
             lty = "dashed") +
  theme(
    axis.text = element_text(size = 12),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 10),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    plot.title = element_text(size = 14, face = "bold"),
    panel.spacing = unit(0, "lines")
  )

# boxplot
box1 <-
  dp %>%
  filter(interval != "PBS", tissue == "spleen", coating == "S2P") %>%
  mutate(interval = factor(interval, levels = c("PBS", 0:8))) %>%
  mutate(
    interval = forcats::fct_recode(
      interval,
      "8 weeks" = "8",
      "6 weeks" = "6",
      "4 weeks" = "4",
      "3 weeks" = "3",
      "2 weeks" = "2",
      "1 week" = "1",
      "Prime only" = "0"
    )
  ) %>%
  mutate(day = forcats::fct_recode(factor(day), "day 64" = "64")) %>%
  ggplot(aes(x = interval, y = counts)) +
  geom_boxplot(outlier.shape = NA) +
  scale_shape_manual(values = c(16, 5)) +
  scale_y_log10(breaks = c(100, 500, 1000, 1500)) +
  geom_jitter(aes(x = interval, y = counts)) +
  stat_summary(
    fun.min = mean,
    fun = mean,
    fun.max = mean,
    color = "red",
    width = .5,
    geom = "errorbar",
    size = 1
  ) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  labs(
    x = "",
    y = "antigen specific \n ASC counts/million cells",
    color = "",
    linetype = "",
    title = "S2P specific spleen day 64 (1wk post-boost) results"
  ) +
  scale_color_manual(values = c(
    "#6496cd",
    "#cb6939",
    "#8264cb",
    "#60a75c",
    "#c361aa",
    "#ad963e",
    "#cc566a"
  )) +
  theme(
    axis.text = element_text(size = 12),
    title = element_text(size = 14),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "bottom"
  )


cowplot::plot_grid(plotlist = list( box1, foldchange1 ), nrow = 2, ncol = 1)
```



### Supplementary Table 3. Statistical comparisons of S2-P–specific ASCs 1 week following dose 2 between mRNA-1273 (10 μg) dosing intervals.
```{r}
#-------------------------------------------------------------------------------
# Supplementary Table 3
#-------------------------------------------------------------------------------
emm_elispot <-
  emmeans(ref_grid(lm_fits$S2P, cov.keep = c("interval", "day")),
          specs = ~ interval,
          type = "response")

set.seed(100)
sum_elispot <-
  contrast(
    regrid(emm_elispot, transform = "log10"),
    "pairwise",
    infer = TRUE,
    adjust = "mvt",
    ratio = FALSE,
    level = .95
  ) %>% as_tibble() %>%
  bind_rows()

sum_elispot  %>%
  as_tibble() %>%
  dplyr::select(contrast, estimate, p.value) %>%
  filter(p.value < 0.05) %>%
  mutate(estimate = 10 ^ estimate) %>%
  mutate(contrast = str_replace(contrast, " - ", " / ")) %>%
  rename(
    "Pairwise comparison" = contrast,
    "Fold change" = "estimate",
    "Adjusted P-value" = "p.value"
  ) %>%
  DT::datatable(caption = "Statistical comparisons of S2-P–specific ASCs 1 week following dose 2 between mRNA-1273 (10 μg) dosing interval") %>%
  formatSignif(columns = c('Fold change', 'Adjusted P-value'),
               digits = 7)


# Supplementary Table 3 visualization
t3v <-
sum_elispot %>%
  as_tibble() %>%
  mutate(contrast = str_replace(contrast, " - ", " / "),
         contrast = str_remove_all(contrast, "interval")) %>%
  mutate(day = "Day 64 (1wk post-boost)")%>%
  rename(
    "Pairwise comparison" = contrast,
    "Fold change" = "estimate",
    "Adjusted P-value" = "p.value",
  ) %>%
  mutate(`Pairwise comparison` = str_remove_all(`Pairwise comparison` , "interval")) %>%
  ggplot(aes(x = `Pairwise comparison` , y = `Fold change`)) +
  geom_point(size = 3)  +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = .1,
                #position = position_dodge(width = 2),
                size = 1) +
  scale_color_manual(
    values = c(
      "#ABCFE5",
      "#93C4DE",
      "#78B5D8",
      "#60A6D1",
      "#4A97C9",
      "#3787C0",
      "#2575B7",
      "#1664AB",
      "#09539D",
      "#084185"
    )
  ) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  scale_y_continuous(breaks = log10(c(0.1, 0.2, 0.5, 1, 2, 3)),
                     labels = c("1/10", "1/5", "1/2", 1, 2, 3)) +
  facet_grid(~day) +
  geom_hline(yintercept = 0,
             color = "darkcyan",
             lty = "dashed") +
  labs(
    title = "Estimated fold change of S2-P–specific ASCs",
    #subtitle = "1 week following dose 2 between mRNA-1273 (10 μg) dosing interval",
    x = "",
    y = ""
  ) +
  coord_flip() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    strip.text = element_text(size = 10),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    plot.title = element_text(size = 14, face = "bold"),
    panel.spacing = unit(0, "lines")
  )


```

## Model fitting for BM data
```{r}
#-------------------------------------------------------------------------------
# Fit the model for BM data 
#-------------------------------------------------------------------------------
# BM 
d_bm <- subset(d, tissue == "BM")

boxcox_fits <-
  d_bm %>%
  split(.$coating) %>%
  map( ~ boxcox_fit(.x))

lm_fits <-
  purrr::transpose(boxcox_fits)$lm_fit

# obtain emmeans
lm_grid <- ref_grid(lm_fits$S2P, cov.keep = c("interval", "day"))
emmt_dat <-
  emmeans(lm_grid, specs = ~ interval | day, type = "response") %>%
  as_tibble()

```

### Residual diagnostics for BM data
```{r, fig.width = 12, fig.height = 6}
assumption_plots <-
  imap(
    lm_fits ,
    ~ assumption_check(.x) %>%
      plot_grid(
        ggdraw() + draw_label(
          paste0("BM ", .y, "-specific model residual diagnostics "),
          x = 0.03,
          hjust = 0
        ),
        .,
        ncol = 1,
        rel_heights = c(0.1, 1),
        align = "v"
      )
  )
ggarrange(plotlist = assumption_plots)
```

### Figure 4B. Spike-specific Antibody Secreting Cells and Long-Lived Plasma Cells

```{r, fig.height=8}
#-------------------------------------------------------------------------------
# Figure 4B. Spike-specific Antibody Secreting Cells and Long-Lived Plasma Cells.
#-------------------------------------------------------------------------------
foldchange2 <-
  emmt_dat %>%
  mutate(
    interval = forcats::fct_recode(
      interval,
      "8 weeks" = "8",
      "6 weeks" = "6",
      "4 weeks" = "4",
      "3 weeks" = "3",
      "2 weeks" = "2",
      "1 week" = "1",
      "Prime only" = "0"
    )
  ) %>%
  mutate(
    day = forcats::fct_recode(
      factor(day),
      "Day 87 (4wk post-boost)" = "87",
      "Day 227 (24wk post-boost)" = "226"
    )
  ) %>%
  mutate(day = factor(day, c(
    "Day 87 (4wk post-boost)" , "Day 227 (24wk post-boost)"
  ))) %>%
  ggplot(aes(x = interval , y = response)) +
  geom_point(position = position_dodge(width = 0.5), size = 2)  +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = .1,
    position = position_dodge(width = 0.5),
    size = 1
  ) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  scale_y_log10(breaks = c(100, 500, 1000, 1600)) +
  ggtitle("S2P specific Bone Marrow results") +
  labs(x = "", y = "Estimated ASC counts/million cells") +
  geom_hline(yintercept = emmt_dat$lower.CL[c(6)],
             color = "darkcyan",
             lty = "dashed") +
  geom_hline(yintercept = emmt_dat$lower.CL[c(13)],
             color = "darkcyan",
             lty = "dashed") +
  facet_grid( ~ day) +
  theme(
    axis.text = element_text(size = 12),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 10),
    legend.text = element_text(size = 14),
    legend.title = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    panel.spacing = unit(0, "lines")
  )

box2 <-
  dp %>%
  filter(interval != "PBS", tissue == "BM", coating == "S2P") %>%
  mutate(interval = factor(interval, levels = c("PBS", 0:8))) %>%
  mutate(
    interval = forcats::fct_recode(
      interval,
      "8 weeks" = "8",
      "6 weeks" = "6",
      "4 weeks" = "4",
      "3 weeks" = "3",
      "2 weeks" = "2",
      "1 week" = "1",
      "Prime only" = "0"
    )
  ) %>%
  mutate(
    day = forcats::fct_recode(
      factor(day),
      "Day 87 (4wk post-boost)" = "87",
      "Day 227 (24wk post-boost)" = "226"
    )
  ) %>%
  mutate(day = factor(
    day,
    levels = c("Day 87 (4wk post-boost)", "Day 227 (24wk post-boost)")
  )) %>%
  ggplot(aes(x = interval, y = counts)) +
  geom_boxplot(outlier.shape = NA) +
  scale_shape_manual(values = c(16, 5)) +
  scale_y_log10(breaks = c(100, 500, 1000, 1600)) +
  geom_jitter(aes(x = interval, y = counts)) +
  stat_summary(
    fun.min = mean,
    fun = mean,
    fun.max = mean,
    color = "red",
    width = .5,
    geom = "errorbar",
    size = 1
  ) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  labs(
    x = "",
    y = "antigen specific \n ASC counts/million cells",
    color = "",
    linetype = "",
    title = "S2P specific Bone Marrow results"
  ) +
  facet_grid( ~ day) +
  theme(
    axis.text = element_text(size = 12),
    title = element_text(size = 14),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "bottom"
  )
```

### Supplementary Table 4. Statistical comparisons of S2-P–specific LLPCs 4 weeks following dose 2 (Day 87) and 24 weeks following dose 2 (Day 227) between mRNA-1273 (10 μg) dosing intervals.
```{r}
#-------------------------------------------------------------------------------
# Supplementary Table 4
#-------------------------------------------------------------------------------
emm_elispot_BM <-
  emmeans(ref_grid(lm_fits$S2P, cov.keep = c("interval", "day")),
          specs = ~ interval |
            day,
          type = "response")

set.seed(100)
emm_elispot_BM <-
  contrast(
    regrid(emm_elispot_BM , transform = "log10"),
    "pairwise",
    infer = TRUE,
    adjust = "mvt",
    ratio = FALSE,
    level = .95
  ) %>% as_tibble() %>%
  bind_rows()

emm_elispot_BM %>%
  as_tibble() %>%
  dplyr::select(day, contrast, estimate, p.value) %>%
  filter(p.value < 0.05) %>%
  mutate(day = as.numeric(as.character(day))) %>%
  arrange(day) %>%
  mutate(estimate = 10 ^ estimate) %>%
  mutate(contrast = str_replace(contrast, " - ", " / ")) %>%
  rename(
    "Pairwise comparison" = contrast,
    "Fold change" = "estimate",
    "Adjusted P-value" = "p.value",
    "Day" = "day"
  ) %>%
  DT::datatable(
    caption = "Statistical comparisons of S2-P–specific LLPCs 4 weeks following
                dose 2 (Day 87) and 24 weeks following dose 2 (Day 227) between mRNA-1273
                (10 μg) dosing intervals"
  ) %>%
  formatSignif(columns = c('Fold change', 'Adjusted P-value'),
               digits = 7)

```


### Supplementary Figure 5. Statistical comparisons of S2-P-specific Antibody Secreting Cells and Long-Lived Plasma Cells between mRNA-1273 (10 µg) dosing intervals(Supplementary Table 3 and 4 visualization)
```{r, fig.width = 16, fig.height=8}
t4v <-
  emm_elispot_BM %>%
  as_tibble() %>%
  mutate(contrast = str_replace(contrast, " - ", " / "),
         contrast = str_remove_all(contrast, "interval")) %>%
  mutate(day = factor(day, levels = c(87, 226))) %>%
  mutate(
    day = forcats::fct_recode(
      day,
      "Day 87 (4wk post-boost)" = "87",
      "Day 227 (24wk post-boost)" = "226",
    )
  ) %>%
  rename(
    "Pairwise comparison" = contrast,
    "Fold change" = "estimate",
    "Adjusted P-value" = "p.value",
  ) %>% 
  mutate(`Pairwise comparison` = str_remove_all(`Pairwise comparison` , "interval")) %>%
  ggplot(aes(x = `Pairwise comparison` , y = `Fold change`)) +
  geom_point(size = 3)  +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = .1,
                #position = position_dodge(width = 2),
                size = 1) +
  scale_color_manual(
    values = c(
      "#ABCFE5",
      "#93C4DE",
      "#78B5D8",
      "#60A6D1",
      "#4A97C9",
      "#3787C0",
      "#2575B7",
      "#1664AB",
      "#09539D",
      "#084185"
    )
  ) +
  theme_bw() +  theme(panel.grid.major = element_line(colour = "grey100")) +
  scale_y_continuous(breaks = log10(c(0.1, 0.2, 0.5, 1, 2, 3)),
                     labels = c("1/10", "1/5", "1/2", 1, 2, 3)) +
  geom_hline(yintercept = 0,
             color = "darkcyan",
             lty = "dashed") +
  facet_grid( ~ day) +
  labs(title = "Estimated fold change of S2-P–specific LLPC",
       # subtitle = "4 weeks following dose 2 (Day 87) and 24 weeks following dose 2 (Day 227)
       # between mRNA-1273 (10 μg) dosing intervals",
       x = "",
       y = "") +
  coord_flip() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    strip.text = element_text(size = 10),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    plot.title = element_text(size = 14, face = "bold"),
    panel.spacing = unit(0, "lines")
  )

cowplot::plot_grid(
  plotlist = list(t3v, t4v),
  labels = c("A", "B"),
  nrow = 1,
  ncol = 2,
  rel_widths = c(0.5, 1)
)
#ggsave(filename = "figures_tables/ELISpot/emmeans_BM_spleen.png", width = 16, height=8)
```

## Session information

```{r}
sessionInfo()
```

